On The Subspace Learning for Network Attack Detection

### 2020-04-28 docs/thesis/chapters/3_mobile/mobile.tex
updating thesis latex

This work proposes an architecture and approach for user behavior analysis based on Eigensim \cite{vieira2017model}, in order to detect possible threats in offline corporate mobile applications. The key expiration period is safely incorporated into the proposed system solution in order to enhance security and the behavioral analysis can indicate malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_eigensim} explains the use of Eigensim. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

\section{Related works}
\label{sec:3_related_works}

@@

Preventing data leakage on the mobile device is a crucial security problem. Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the mobile client or the business itself?

@@

The necessity to use schemes that function without putting load on a provider arises when it is desired to make the mobile client less dependent on the cloud provider, i.e. corporate client continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable, heydari2016efficient}, a combination of both \cite{itani2010energy} or ad hoc approaches \cite{shila2016amcloud}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}. Therefore, according to \cite{khan2013towards, khan2015cloud}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.

To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The mobile security architecture}
\label{sec:3_the_client_security_architecture}

@@

The core set of functions and protocols of the proposed architecture can be divided into three sets of operations as shown in Figure \ref{fig:3_01}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/ch3/fig01.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:3_01}
\end{figure}

Figure \ref{fig:3_01} describes the mobile client protection both in online and offline mode. In online mode, the client has the possibility to connect to the server and the security of the client is enhanced by the server-backed up mechanisms. On the other hand, in offline mode the client’s security is supported by the standalone mechanisms. Additionally, the mobile client protection is enhanced by the threat intelligence unit providing the constant monitoring and analysis.

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: 
\begin{enumerate}
	\item Encryption subsystem that provides the procedures of encryption and decryption;
	\item Protected storage subsystem that provides the downloaded shares and key storage protection;
	\item The Threat intelligence unit that provides the constant monitoring;
	\item The communication subsystem that enables with the server.
\end{enumerate}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/ch3/fig02.eps}
	\caption{The Client-side Architecture}
	\label{fig:3_02}
\end{figure}

@@

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and Eigensim-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this module as well. These functions are described in Subsection \ref{sec:3_eigensim}.

\section{The proposed solution for offline mobile security}
\label{sec:3_offline_mode}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=15cm]{figures/ch3/fig03.eps}
	\caption{Proposed Architecture for Offline Mobile Security}
	\label{fig:3_03}
\end{figure}

The Figure \ref{fig:3_03} describes the proposed architecture for offline mobile security, showing the modules responsible for securing the mobile client, which includes:

\begin{enumerate}
	\item \pmb{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
	\item \pmb{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal,vieira2017model} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec:3_results}. Moreover, the Eigensim, which is an approach based on subspace learning and on effective signal processing technique to separate noise components from the principal components, named Model Order Selection (MOS), can be applied into anomaly and attack detection \cite{tenorio2013greatest, vieira2017model}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \pmb{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

\subsection{Offline Behavioral Analysis}
\label{sec:3_offline_behavioral_analysis}

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:3_06} depicts the TIM workflow for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=10cm]{figures/ch3/fig06.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:3_06}
\end{figure}

As depicted in Figure \ref{fig:3_06}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes proposed by Eigensim, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec:3_eigensim}.

\section{Eigensim for Threat Intelligence}
\label{sec:3_eigensim}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as subspace learning by eigenvalue decomposition and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors. Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.

Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, such as user operations, should be modeled as a matrix composed by the number of occurrences grouped by location and time, and splited into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\pmb{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.

The Eigensim can rely on sample covariance of zero mean variables (called as \pmb{zero mean covariance} for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as \pmb{zero mean and standardized covariance} for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

For Eigensim based on sample covariance of zero mean and unitary standard deviation, in order to identify anomalies with sparse or subtle behavior, it is required, for each variable, to make the standard deviation unitary as follows

@@

Once the $\hat{\pmb{R}}_{yy}$ or $\hat{\pmb{R}}_{zz}$ have been obtained for anomaly detection based on Eigensim, for the sake of simplicity, we refer to $\hat{\pmb{R}}_{yy}$ or $\hat{\pmb{R}}_{zz}$ as a matrix $\hat{\pmb{R}}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\hat{\pmb{R}}^{(q)} = \pmb{V}^{(q)}\pmb{\lambda}^{(q)}\pmb{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $\pmb{e}$, as following:

@@

Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\pmb{e}_{\rm max}, M)$. For more information about MOS schemes, interested readers are referred to Appendix \ref{apx:a_mos} and to \cite{da2009comparison}.

\section{Results and analysis}
\label{sec:3_results}

@@

\subsection{Security analysis}
\label{sec:3_sec_analysis}
The security analysis of the proposed model was performed from the user behavioral analysis. Two common attack scenarios were analyzed. First, the malicious outsider trying to infect or steal the important files. Second, the malicious expired user trying to steal the important files.

\subsubsection{Common threat scenarios}
\label{sec:3_common}
This section provides the detailed description of the common scenarios in which the log and behavioral analysis is provided. The behavioral analysis can help to keep the user or administrator informed of the threat and take actions, as well as it can be useful in order to implement threat preventions or reactive actions to avoid threat propagation.

@@

The Eigensim and MOS schemes based on zero mean covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.

@@

Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. Eigensim and zero mean and standardized covariance analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation

The Eigensim based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. Subsequently, the eigenvalue analysis highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The session time is valid;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time.
\end{enumerate}

@@

In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile device, to emulate a period in which the session was valid. The log analysis by Eigensim can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.

@@

Applying Eigensim to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.

The Eigensim based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the zero mean covariance matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals the operating system;
	\item The hacker modifies the time of the operating system to a period when the session was valid;
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The mobile client verifies if one logged time is older than the last operation time. If it is true, the Eigensim module classifies the evaluated operation as malicious;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time;
\end{enumerate}

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec:3_data}

The Eigensim and MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to Eigensim, according to described in Section \ref{sec:3_eigensim}.

The selected features shall be modeled as matrices which represents a signal superposition of signal, artifact and noise, which refer to normal, controlled and anomalous behaviors \cite{tenorio2013greatest,vieira2017model}. Thus, the data is grouped into time frames $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.

@@

\subsection{Performance analysis}
\label{sec:3_complex}

@@

Table \ref{tab:3_04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.

>>>>

In this chapter we propose an architecture and approach for user behavior analysis based on Eigensimilarity \citeme{vieira2017model}, in order to detect possible threats in offline corporate mobile applications \citeme{galibus2016organization, galibus2017offline}. The key expiration period is safely incorporated into the proposed system solution in order to enhance security and the behavioral analysis can indicate malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. In Section \ref{sec:a_related_works} we analyze the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. The mobile security architecture of the proposed solution is outlined in Section \ref{sec:a_the_client_security_architecture}, and in Section \ref{sec:a_offline_mode} is presented the detailed scheme of the proposed solution to the problem of offline mobile client security. In Section \ref{sec:a_Eigensimilarity} we explain the use of Eigensimilarity and in Section \ref{sec:a_results} we discuss the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results. Finally, in Section \ref{sec:a_conclusion} the chapter is concluded.

\section{Related works}
\label{sec:a_related_works}

@@

Preventing data leakage on the mobile device is a crucial security problem. Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME needs to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the mobile client or the business itself?

@@

The necessity to use schemes that function without putting load on a provider arises when it is desired to make the mobile client less dependent on the cloud provider, i.e. corporate client continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable, heydari2016efficient}, a combination of both \cite{itani2010energy} or \emph{ad hoc} approaches \cite{shila2016amcloud}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}. Therefore, according to \cite{khan2013towards, khan2015cloud}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.

To the best of our knowledge, the offline mode security problem has not been deployed yet, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The mobile security architecture}
\label{sec:a_the_client_security_architecture}

@@

The core set of functions and protocols of the proposed architecture can be divided into three sets of operations as shown in Figure \ref{fig:a_01}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/ch3/fig01.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:a_01}
\end{figure}

Figure \ref{fig:a_01} describes the mobile client protection both in online and offline mode. In online mode, the client has the possibility to connect to the server and the security of the client is enhanced by the server-backed up mechanisms. On the other hand, in offline mode the client’s security is supported by the standalone mechanisms. Additionally, the mobile client protection is enhanced by the threat intelligence unit providing the constant monitoring and analysis.

Figure \ref{fig:a_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: 
\begin{enumerate}
	\item Encryption subsystem that provides the procedures of encryption and decryption;
	\item Protected storage subsystem that provides the downloaded shares and key storage protection;
	\item The Threat intelligence unit that provides the constant monitoring;
	\item The communication subsystem that enables with the server.
\end{enumerate}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/ch3/fig02.eps}
	\caption{The Client-side Architecture}
	\label{fig:a_02}
\end{figure}

@@

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and Eigensimilarity-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this module as well. These functions are described in Subsection \ref{sec:a_Eigensimilarity}.

\section{The proposed solution for offline mobile security}
\label{sec:a_offline_mode}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=15cm]{figures/ch3/fig03.eps}
	\caption{Proposed Architecture for Offline Mobile Security}
	\label{fig:a_03}
\end{figure}

The Figure \ref{fig:a_03} describes the proposed architecture for offline mobile security, showing the modules responsible for securing the mobile client, which includes:

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec:a_results}. Moreover, the Eigensimilarity, which is an approach based on subspace learning and on effective signal processing technique to separate noise components from the principal components \cite{lyra2012improved, gomes2019tensor}, named Model Order Selection (MOS), can be applied into anomaly and attack detection \citeme{vieira2017model}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

\subsection{Offline Behavioral Analysis}
\label{sec:a_offline_behavioral_analysis}

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:a_06} depicts the TIM workflow for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=10cm]{figures/ch3/fig06.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:a_06}
\end{figure}

As depicted in Figure \ref{fig:a_06}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes proposed by Eigensimilarity, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec:a_Eigensimilarity}.

\section{Eigensimilarity for Threat Intelligence}
\label{sec:a_Eigensimilarity}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as subspace learning by eigenvalue decomposition and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors. Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.

Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, such as user operations, should be modeled as a matrix composed by the number of occurrences grouped by location and time, and split into $Q$ time frames. Therefore, the framework considers the time variations of the matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors. For example, one element of $\pmb{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.

The Eigensimilarity can rely on sample covariance of zero mean variables (called as \textbf{zero mean covariance} for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as \textbf{zero mean and standardized covariance} for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.

@@

For Eigensimilarity based on sample covariance of zero mean and unitary standard deviation, in order to identify anomalies with sparse or subtle behavior, it is required, for each variable, to make the standard deviation unitary as follows

@@

Once the $\hat{\pmb{R}}_{yy}$ or $\hat{\pmb{R}}_{zz}$ have been obtained for anomaly detection based on Eigensimilarity, for the sake of simplicity, we refer to $\hat{\pmb{R}}_{yy}$ or $\hat{\pmb{R}}_{zz}$ as a matrix $\hat{\pmb{R}}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\hat{\pmb{R}}^{(q)} = \pmb{V}^{(q)}\pmb{\lambda}^{(q)}\pmb{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $\pmb{e}$, as following:

@@

Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\pmb{e}_{\rm max}, M)$. For more information about MOS schemes, interested readers are referred to section \ref{sec:2_mos} and to \cite{da2009comparison}.

\section{Results and analysis}
\label{sec:a_results}

@@

\subsection{Security analysis}
\label{sec:a_sec_analysis}
The security analysis of the proposed model was performed from the user behavioral analysis. Two common attack scenarios were analyzed. First, the malicious outsider trying to infect or steal the important files. Second, the malicious expired user trying to steal the important files.

\subsubsection{Common threat scenarios}
\label{sec:a_common}
This section provides the detailed description of the common scenarios in which the log and behavioral analysis is provided. The behavioral analysis can help to keep the user or administrator informed of the threat and take actions, as well as it can be useful in order to implement threat prevention or reactive actions to avoid threat propagation.

@@

The Eigensimilarity and MOS schemes based on zero mean covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.

@@

Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. Eigensimilarity and zero mean and standardized covariance analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation

The Eigensimilarity based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. Subsequently, the eigenvalue analysis highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The session time is valid;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigation;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time.
\end{enumerate}

@@

In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile device, to emulate a period in which the session was valid. The log analysis by Eigensimilarity can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.

@@

Applying Eigensimilarity to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.

The Eigensimilarity based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the zero mean covariance matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.

@@

\begin{enumerate}[label=(\alph*)]
	\item The hacker steals the operating system;
	\item The hacker modifies the time of the operating system to a period when the session was valid;
	\item The hacker has access to the mobile client and is able to perform operations;
	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
	\item The mobile client incrementally append each operation attempt time into the logging;
	\item The mobile client verifies if one logged time is older than the last operation time. If it is true, the Eigensimilarity module classifies the evaluated operation as malicious;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
	\item The TIM module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time;
\end{enumerate}

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec:a_data}

The Eigensimilarity and MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to Eigensimilarity, according to described in Section \ref{sec:a_Eigensimilarity}.

Thus, the data is grouped into time frames $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.

@@

\subsection{Performance analysis}
\label{sec:a_complex}

@@

Table \ref{tab:a_04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.

\begin{table*}
	\caption{Data Modeling and Eigenvalue Decomposition Time}
	\scriptsize
	\label{tab:a_04}
	\centering
	\begin{tabular}{|l|l|l|l|l|l|l|l|}
		\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Modeling\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg. Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std. Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig. Min.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig. Max.\\(ms)\end{tabular}	\\ \hline
		Galaxy GT-I9300	& 6	& 60	& 107	& 209.52	& 18.58	& 183	& 276	\\ \hline
		Galaxy GT-I9300	& 6	& 40	& 115	& 227.26	& 18.13	& 191	& 289	\\ \hline
		Galaxy GT-I9300	& 6	& 20	& 89	& 268.14	& 21.94	& 229	& 315	\\ \hline
		Galaxy GT-I9300	& 6	& 10	& 90	& 347.42	& 24.11	& 304	& 421	\\ \hline
		Galaxy GT-I9300	& 4.1	& 60	& 20	& 60.90	& 15.19	& 37	& 106	\\ \hline
		Galaxy GT-I9300	& 4.1	& 40	& 20	& 68.72	& 15.71	& 43	& 114	\\ \hline
		Galaxy GT-I9300	& 4.1	& 20	& 34	& 89.04	& 16.78	& 54	& 133	\\ \hline
		Galaxy GT-I9300	& 4.1	& 10	& 21	& 117.24	& 14.36	& 96	& 171	\\ \hline
		Galaxy GT-I9300	& 1.4	& 60	& 10	& 159.82	& 15.82	& 125	& 197	\\ \hline
		Galaxy GT-I9300	& 1.4	& 40	& 10	& 168.06	& 15.90	& 139	& 220	\\ \hline
		Galaxy GT-I9300	& 1.4	& 20	& 11	& 204.4	& 20.46	& 176	& 269	\\ \hline
		Galaxy GT-I9300	& 1.4	& 10	& 13	& 259.00	& 21.34	& 220	& 315	\\ \hline
		Galaxy Tab SM-T800	& 6	& 60	& 7	& 59.30	& 6.55	& 54	& 74	\\ \hline
		Galaxy Tab SM-T800	& 6	& 40	& 8	& 62.56	& 7.05	& 56	& 80	\\ \hline
		Galaxy Tab SM-T800	& 6	& 20	& 10	& 73.28	& 8.59	& 65	& 95	\\ \hline
		Galaxy Tab SM-T800	& 6	& 10	& 8	& 93.48	& 9.13	& 83	& 130	\\ \hline
		Galaxy Tab SM-T800	& 4.1	& 60	& 11	& 18.64	& 4.51	& 16	& 38	\\ \hline
		Galaxy Tab SM-T800	& 4.1	& 40	& 11	& 19.64	& 5.12	& 17	& 38	\\ \hline
		Galaxy Tab SM-T800	& 4.1	& 20	& 12	& 25.12	& 5.55	& 21	& 46	\\ \hline
		Galaxy Tab SM-T800	& 4.1	& 10	& 12	& 32.32	& 7.29	& 27	& 55	\\ \hline
		Galaxy Tab SM-T800	& 1.4	& 60	& 4	& 49.08	& 6.01	& 42	& 62	\\ \hline
		Galaxy Tab SM-T800	& 1.4	& 40	& 5	& 51.42	& 7.36	& 44	& 74	\\ \hline
		Galaxy Tab SM-T800	& 1.4	& 20	& 5	& 51.12	& 7.80	& 54	& 91	\\ \hline
		Galaxy Tab SM-T800	& 1.4	& 10	& 7	& 75.24	& 7.71	& 65	& 90	\\ \hline
	\end{tabular}
\end{table*}
